---
title: "Homework 2"
output:
  html_document:
    theme: yeti
    highlight: haddock
---

```{r}
library(tidyverse)
library(readxl)
library(stringr) 
```

# Problem 1

### Cleaning trash_wheel Data

```{r}
trash_wheel = read_excel("data/trash_wheel.xlsx", range = "A2:N406") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0)) %>% 
  select("dumpster", "homes_powered")
```

### Cleaning precipitation data 2019 

```{r}
rain_2019 = read_excel("data/trash_wheel.xlsx", 
  sheet = "2019 Precipitation", skip = 1)  %>%
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = "2019")
```

### Cleaning precipitation data 2018

```{r}
rain_2018 = read_excel("data/trash_wheel.xlsx", 
  sheet = "2018 Precipitation", skip = 1)  %>%
  janitor::clean_names() %>% 
  drop_na(month)  %>% 
  mutate(year = "2018")
```

### Combining the precipitation datasets

```{r}
combined_rain = 
  bind_rows(rain_2018, rain_2019) %>% 
  mutate(month = month.name[as.numeric(month)])
```

## Description of datasets

### Trash Wheel Dataset

* The Trash Wheel dataset includes `r length(trash_wheel)` variables: `r variable.names(trash_wheel)`.
* The number of observations in the Trash Wheel dataset is `r count(trash_wheel)`.
* The dataset has `r nrow(trash_wheel)` rows and `r ncol(trash_wheel)` columns.
* The range of the sports balls is (`r range(trash_wheel$sports_balls)`) mm.
* The median number of sports balls is (`r median(trash_wheel$sports_balls, year = 2019)`).

### Precipitation Dataset

* The combined precipitation dataset includes `r length(combined_rain)` variables: `r variable.names(combined_rain)`.
* The number of observations in the combined precipitation  dataset is `r count(combined_rain)`.
* The dataset has `r nrow(combined_rain)` rows and `r ncol(combined_rain)` columns.
* The range of precipitation is (`r range(combined_rain$total)`).
* The total precipitation in 2018 was `r sum(rain_2018$total)`.

# Problem 2

### Cleaning Pols_months data

```{r}
pols_months =  
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    prez_dem = recode(prez_dem, `1` = "dem", `0` = "gop"),
    prez_dem = factor(prez_dem),
    year = as.numeric(year),
    month = str_remove(month, "^0+"),  
    month = as.numeric(month)) %>% 
  rename(president = prez_dem) %>% 
  select(-prez_gop, -day) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(month = str_to_lower(month)) 
```

### Cleaning Snp data

```{r}
snp = 
  read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  relocate(year, month, everything()) %>% 
  select(-day) %>% 
  mutate(
    year = as.numeric(year),
    month = month.name[as.numeric(month)]) %>% 
  mutate( 
   year = ifelse(year >= 21, year+1900, year+2000),
   month = str_to_lower(month)) 
```

### Cleaning Unemployment data

```{r}
unemployment = 
  read_csv("data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "close"
  )
```

### Joining datasets

```{r}
snp_pols_combined = 
 full_join(pols_months, snp, by = c("year", "month"))
final_combined = 
 full_join(snp_pols_combined, unemployment, by = c("year", "month"))
```

### Description of datasets

#### Pols Dataset Description 

* The Pols dataset includes `r length(pols_months)` variables: `r variable.names(pols_months)`.
* The number of observations in the Pols dataset is `r count(pols_months)`.
* The dataset has `r nrow(pols_months)` rows and `r ncol(pols_months)` columns.
* The range of the years is (`r range(pols_months$year)`).

#### Snp Dataset Description 

* The Pols dataset includes `r length(snp)` variables: `r variable.names(snp)`.
* The number of observations in the Snp dataset is `r count(snp)`.
* The dataset has `r nrow(snp)` rows and `r ncol(snp)` columns.
* The range of the years is (`r range(snp$year)`).

#### Unemployment Dataset Description 

* The Pols dataset includes `r length(unemployment)` variables: `r variable.names(unemployment)`.
* The number of observations in the Unemployment dataset is `r count(unemployment)`.
* The dataset has `r nrow(unemployment)` rows and `r ncol(unemployment)` columns.
* The range of the years is (`r range(unemployment$year)`).

#### Combined Dataset Description 

* The combined dataset includes `r length(final_combined)` variables: `r variable.names(final_combined)`.
* The number of observations in the combined dataset is `r count(final_combined)`.
* The dataset has `r nrow(final_combined)` rows and `r ncol(final_combined)` columns.
* The range of the years is (`r range(final_combined$year)`).

# Problem 3

## Cleaning baby names dataset

```{r}
baby_names = 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender)) %>% 
  mutate(ethnicity = str_to_lower(ethnicity)) %>%  
  mutate(childs_first_name = str_to_lower(childs_first_name)) %>% 
  distinct(year_of_birth, gender, ethnicity, childs_first_name, .keep_all = TRUE) 
```

### Olivia Table 

```{r}
olivia_popularity = 
 filter(baby_names, childs_first_name == "olivia") %>% 
 select(-gender, -count, -childs_first_name)  
```

```{r}
knitr::kable(head(olivia_popularity[, 1:3]), "simple")
```

### Jayden Table 

```{r}
jayden_popularity = 
 filter(baby_names, childs_first_name == "jayden") %>% 
 select(-gender, -count, -childs_first_name)  
```
 
```{r}
knitr::kable(head(jayden_popularity[, 1:3]), "simple")
```

### Scatterplot

```{r}
scatterplot_data =   
  filter(baby_names, gender == "male", ethnicity == "white non hispanic", year_of_birth == 2016)
```

```{r}
ggplot(scatterplot_data, aes(x = rank, y = count)) + 
    geom_point()
```


